Model Search in Relational Domains

### 2016-10-19 experiments.tex
comments in section 7

We verified a total of 66 benchmarks in ANSI-C, which are derived from 
(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
cycle accurate software models of hardware circuits automatically 
generated using v2c~\cite{mtk2016}, (c) controller code generated 
from Simulink (d) several hand-crafted benchmarks for equivalence 
checking and bounded loop analysis.  All bounded loops are completely 
unrolled before analysis.  We compare our tool with a state-of-the-art 
SAT-based bounded model checker CBMC~\cite{cbmc} and a commercial 
static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
Astr{\'e}e was configured with default set of domains which includes 
interval, bit-field, congurence, trace partitioning and relational 
domains (octagons, polyhedra, zones, equalities, filter).

\Omit {
To enable precise analysis using Astr{\'e}e, all our benchmarks are 
manually instrumented with partition directives which provides external 
hint to the tool to guide the trace partitioning heuristics.  Usually, 
such high-precision is not needed for static analysis, since it makes 
the analysis very expensive.  Without trace partitioning, the 
analysis using Astr{\'e}e shows high degree of imprecision. 
}
\paragraph {\textbf{Precision and Efficiency of Analysis}}
Figure~\ref{fig:results} presents a comparison of the solver statistics
(decision and learning) for the analysis using CBMC and ACDCL.  The cumulative 
statistics corresponding to each phase of the algorithm for all 66 benchmarks 
is presented in Table~\ref{result}.  The reasoning using ACDCL is atleast 
\rmcmt{1.5X} effecient compared to SAT in terms of number of decisions, propagations 
and learning.    
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{result}
\end{table}
%
Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16 benchmarks 
without any restarts.  The solver is restarted in 50 cases to avoid spending 
too much time in branches that does not easily lead to a satisfying assignment or 
stronger clause learning.  In contrast, the analysis using ACDCL solved all
66 benchmarks without any restarts.  This is attributed to intelligent decision 
heuristics which exploits the high-level structure of the program combined 
with stronger deductions and clause learning mechanism aided by the richer 
abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
but the analysis using Astr{\'e}e shows high degree of imprecision.  
Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.

\paragraph {\textbf{Decision Heuristics}} We compare the performance of 
different decision heuristics in ACDCL.  We observe that the ordered heuristic 
outperforms other heuristics on control-intensive benchmarks due to its 
ability to prioritize decisions on variables that appears in conditional 
branches.  For straight-line code and most of the bit-vector regression 
suite, random heuristics performs the best.  Whereas, the activity based 
heuristics such as Berkmin heuristic which works well in propositional 
cases performs best for benchmarks that encountered the maximum number 
of conflicts to prove safety, thus allowing the heuristics to choose 
the decison variable among the set of learnt clauses.

\paragraph {\textbf{Propagation Strategy}}      
Figure~\ref{prop-dec} presents a comparison between {\em forward} analysis and
{\em chaotic} analysis in ACDCL.  The choice of analysis influences the 
total number of decisions and clause learning iterations.  Hence, the 
propagation strategy has a significant influence on the run time which 
can be visualised in Figure~\ref{prop-dec}.  Unlike the forward analysis, 
the chaotic or multi-way strategy performs both forward and backward 
propagations.  This may take several iterations to reach the fixed-point, but 
significantly reduces the total number of decisions and learning.  
\Omit {
We observe that the deductions made during chaotic analysis helps the decision 
heuristics to make better decisions compared to forward analysis.  This is 
attributed to the precision of deductions inferred by the chaotic analysis.         
}

\paragraph {\textbf{Learning}} Learning has a significant influence on
the runtime of ACDCL.  We compare the first-UIP learning technique with 
a DPLL-style conflict analysis without learning.  The effect of UIP computation 
allows ACDCL to backtrack non-chronologically and guide the model search with 
a learnt transformer.  However, the DPLL-style analysis exhibits
case-enumeration behavior and thus immediately runs out of capacity for programs 
which require relational domains.

>>>>

We verified a total of 66 benchmarks in ANSI-C, which are derived from 
(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
cycle accurate software models of hardware circuits automatically 
generated using v2c~\cite{mtk2016}, (c) controller code generated 
from Simulink (d) several hand-crafted benchmarks for equivalence 
checking and bounded loop analysis.  All bounded loops are completely 
unrolled before analysis.  We compare our tool with a state-of-the-art 
SAT-based bounded model checker CBMC~\cite{cbmc} and a commercial 
static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
Astr{\'e}e was configured with default set of domains which includes 
interval, bit-field, congruence, trace partitioning and relational 
domains (octagons, polyhedra, zones, equalities, filter).

\Omit {
To enable precise analysis using Astr{\'e}e, all our benchmarks are 
manually instrumented with partition directives which provides external 
hint to the tool to guide the trace partitioning heuristics.  Usually, 
such high-precision is not needed for static analysis, since it makes 
the analysis very expensive.  Without trace partitioning, the 
analysis using Astr{\'e}e shows high degree of imprecision. 
}
\paragraph {\textbf{Precision and Efficiency of Analysis}}
Figure~\ref{fig:results} presents a comparison of the solver statistics
(decision and learning) for the analysis using CBMC and ACDCL.  The cumulative 
statistics corresponding to each phase of the algorithm for all 66 benchmarks 
is presented in Table~\ref{result}.  The reasoning using ACDCL is at least 
\rmcmt{1.5X} more efficient compared to propositional SAT in terms of number of decisions, propagations 
and learning \pscmt{what learning???}.    
%
\begin{table}
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{result}
\end{table}
%
Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16 benchmarks 
without any restarts.  The solver is restarted in 50 cases to avoid spending 
too much time in branches that does not easily lead to a satisfying assignment or 
stronger clause learning.  In contrast, the analysis using ACDCL solved all
66 benchmarks without any restarts.  This is attributed to intelligent decision 
heuristics which exploits the high-level structure of the program combined 
with stronger deductions and clause learning mechanism aided by the richer 
abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
but the analysis using Astr{\'e}e shows high degree of imprecision.  
Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.

\paragraph {\textbf{Decision Heuristics}} We compare the performance of 
different decision heuristics in ACDCL.  We observe that the ordered heuristic 
outperforms other heuristics on control-intensive benchmarks due to its 
ability to prioritize decisions on variables that appear in conditional 
branches.  For straight-line code and most of the bit-vector regression 
suite, random heuristics performs the best.  Whereas activity-based 
heuristics such as Berkmin heuristic which works well in propositional 
cases performs best for benchmarks that encountered \pscmt{???} the maximum number 
of conflicts to prove safety, thus allowing the heuristics \pscmt{???} to choose  
the decison variable among the set of learnt clauses.

\paragraph {\textbf{Propagation Strategy}}      
Figure~\ref{prop-dec} presents a comparison between {\em forward} analysis \pscmt{iteration/propagation strategy} and
{\em chaotic} analysis in ACDCL \pscmt{what about backward?}.  The choice of analysis influences the 
total number of decisions and clause learning iterations.  Hence, the 
propagation strategy has a significant influence on the run time which 
can be visualised in Figure~\ref{prop-dec}. \pscmt{Are the graphs up-to-date?} Unlike the forward analysis, 
the chaotic or multi-way strategy performs both forward and backward 
propagations.  This may take several iterations to reach the fixed-point, but 
significantly reduces the total number of decisions and learning. \pscmt{Why?} 
\Omit {
We observe that the deductions made during chaotic analysis helps the decision 
heuristics to make better decisions compared to forward analysis.  This is 
attributed to the precision of deductions inferred by the chaotic analysis.         
}

\paragraph {\textbf{Learning}} Learning has a significant influence on
the runtime of ACDCL.  We compare the first-UIP learning technique with 
a DPLL-style conflict analysis without learning \pscmt{you learn the negation of the last decision}.  The effect of UIP computation 
allows ACDCL to backtrack non-chronologically and guide the model search with 
a learnt transformer.  However, the DPLL-style analysis exhibits
case-enumeration behavior and thus immediately runs out of capacity for programs 
which require relational domains. \pscmt{Is there a significant difference between intervals and octagons that indicates that for relational domains DPLL is even worse than in the propositional case?}
